Application of Machine Learning to Increase the Efficiency of the Global Search Algorithm for Solving Multicriterial Problems\thanks{This research was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2023-10-16 2023/NUMTA/Full paper/NUMTA_2023_MCO.tex
fix numta

For efficiency confirmation of the proposed approach combining the global optimization algorithm with machine learning procedures, the representative computational experiment has been conducted on the test class of multiextremal multicriterial problems of different dimensions. For expanding the confirmation area, the test problems were solved by several known meta-heuristic methods of multicriterial optimization. The Hyper-Volume (HV) index reflecting the quality of the Pareto set approximation was used as a comparison criterion. The results of the experiment have demonstrated significant advantages of the new algorithm over other methods having involved in comparison.

@@

In decision making terminology the region (\ref{eq:01}) is called \textit{feasible domain}, the argument $y$ belonging to the domain (\ref{eq:01}) is \textit{feasible decision}.  Functions  $\varphi_s(y)$, $1 \leq s \leq k$, are called \textit{partial criteria} of the problem. These functions can be interpreted as goals of decision making and, for certainty, we will assume that the less the function value, the better the result of achieving this goal.

Partial criteria form a vector-function called \textit{vector criterion}
\begin{equation}
\label{eq:02}
    \Phi(y) = (\varphi_1(y), \varphi_2(y), \dots, \varphi_k(y))
\end{equation}
and problem of multicriterial optimization is formulated as the problem of minimizing the vector function (\ref{eq:02}) over the domain (\ref{eq:01}):

@@

If one or more partial criteria are not positive, this is not too restrictive, because it is possible to convert this problem to the statement satisfying conditions (\ref{eq:07}) with preserving the Pareto solution.

>>>>

For efficiency confirmation of the proposed approach combining the global optimization algorithm with machine learning procedures, the representative computational experiment has been conducted on the test class of multiextremal multicriterial problems of different dimensions. For expanding the confirmation area, the test problems were solved by several known meta-heuristic methods of multicriterial optimization. The Hyper-Volume (HV) index reflecting the quality of the Pareto set approximation was used as a comparison criterion. The results of the experiment have demonstrated significant advantages of the new algorithm over other methods.

@@

In decision making terminology the region (\ref{eq:01}) is called \textit{feasible domain}, the argument $y$ belonging to the domain (\ref{eq:01}) is \textit{feasible decision}.  Functions  $\varphi_s(y)$, $1 \leq s \leq k$, are called \textit{partial criteria} of the problem. These functions can be interpreted as goals of decision making and, for certainty, we will assume that the less the function value, the better the result of achieving this goal. In other words, the best result for the $s$-th criterion corresponds to the global minimum of the function $\varphi_s(y)$ in the domain (\ref{eq:01}).

Partial criteria form a vector-function called \textit{vector criterion}. 
\begin{equation}
\label{eq:02}
    \Phi(y) = (\varphi_1(y), \varphi_2(y), \dots, \varphi_k(y))
\end{equation}
and problem of multicriterial optimization is formulated as the problem of minimizing the vector function (\ref{eq:02}) over the domain (\ref{eq:01}):

@@

If one or more partial criteria are not positive, this is not too restrictive, because it is possible to convert this problem to the statement satisfying conditions (\ref{eq:07}) with preserving the Pareto solution \cite{Pardalos2017}.
